Synthesizing Verified Components for Cyber Assured Systems Engineering \thanks{ Funded by DARPA Cyber Assured Systems Engineering (CASE) contract HR00111890001. }

### 2021-05-06 paper/example.tex
Simplify example (fix bug)

The goal is to cyber-harden the system so that it is resilient to cyber-attack. For this simple example, two attacks are considered; both attacks originating from the untrusted route-planner: first, malformed messages to disrupt the waypoint manager or UAV flight control; and second, a delayed or unrequested automation response to prevent the UAV from flying a mission or to fly the wrong mission.

@@

eq nothing : bool = (not req) -> (not pre (req)) and (not req);
eq today : bool = req and rsp;    
eq yesterday : bool = false -> pre (req and not rsp);
eq policy : bool = nothing
                or (today and not yesterday)
                or (not today and rsp and yesterday);
eq since : bool = (alrt and (false -> pre(since))) 
               or (not policy and alrt);

guarantee "Start includes a waypoint" :
    rsp => wp;
guarantee "Locations required after the start waypoint" :
    (wp and not rsp) => avl;
guarantee "Waypoint is well-formed" :
    wp => WELL_FORMED_WAYPOINT(Waypoint);
guarantee "Alert if start is not bounded relative to a request" :
    (policy and not alrt) or since;
\end{lstlisting}
\end{lrbox}

@@

The contract in \figref{fig:sw} uses \texttt{eq} statements to define variables local to the contract. For example, the \texttt{req} variable is equivalent to the \texttt{event(req)} expression. An \texttt{event} expression is true if there is data on the named AADL event port. The system contract assumes well-formed input and guarantees a set of properties about the output.

@@

The guarantee is an invariant on the expression \texttt{(policy and not alrt) or since} meaning that either the policy holds and there is no alert or the alert has been sounding since the policy was first violated. The \texttt{policy} is satisfied if there is no pending request (\texttt{nothing}), a request with a response right now (\texttt{today}), or a request yesterday with its response today. The value of \texttt{nothing} and \texttt{yesterday} in the current time step rely on values from the previous time step. The \texttt{yesterday} uses the \texttt{->} operator for initialization. The left operand is the initial value of \texttt{yesterday} at the start of the system, which in this example is \texttt{false} because there is no yesterday at the start.  The right operand is the value of \texttt{yesterday} after the initial step. Here the \texttt{pre} operator refers to the value of \texttt{(req and not rsp)} in the prior time step. Intuitively, \texttt{yesterday} is true if the previous time step made a request without a matching response.

The value of \texttt{since} relies on its own value in the previous time step. The intuitive reading of the expression is that the alert has been true since the time when the policy first failed. The first failure sets \texttt{since} to true (i.e., \texttt{(not policy and alrt)}), and once \texttt{since} true, it persists as long as \texttt{alrt} holds. As such, the \emph{Alert if start is not bounded relative to a request} guarantee defines one behavior of a correctly implemented system. In this way, the contracts model the expected input and output of the system as a whole, and they also model the expected behavior of each component used to implement the system.

Model checking proves whether the contract models of the components, with their connections, implement the system model in \figref{fig:sw}. The results of the model checker are in \figref{fig:example}(b). Not unexpectedly, since the model of the route planner leaves the output behavior unconstrained (i.e., the component can do anything since it is not trusted) the model checker proves that the component composition does not implement the system requirements.

@@

\newsavebox{\mntr}
\begin{lrbox}{\mntr}
\begin{lstlisting}[style=agree]
const is_latched : bool = 
Get_Property(this, CASE_Properties::Monitor_Latched);
eq rsp : bool = event(Response);
eq req : bool = event(Request);
eq nothing : bool = (not req) -> (not pre (req)) and (not req);
eq today : bool = req and rsp;    
eq yesterday : bool = false -> pre (req and not rsp);
eq policy: bool = nothing
               or (today and not yesterday)
               or (not today and rsp and yesterday);
eq alert : bool = (not policy) 
               -> ((is_latched and pre(alert)) or not policy);
guarantee Monitor_Alert 
  "Alert port tracks alert variable" :
  event(Alert) = alert;
guarantee Monitor_Output
  "Output if not alerted" :
  if event(Alert) then (not event(Output)) else
  if event(Response) then (event(Output) and (Output = Response))
  else (not event(Output));
\end{lstlisting}
\end{lrbox}

>>>>

The goal is to cyber-harden the system so that it is resilient to cyber-attack. For this simple example, two attacks are considered; both attacks originating from the untrusted route-planner: first, malformed messages to disrupt the waypoint manager or UAV flight controller; and second, a delayed or unrequested automation response to prevent the UAV from flying a mission or to fly the wrong mission.

@@

eq today : bool = (req = rsp);
eq yesterday : bool = (req and not rsp) ->
                      pre(req and not rsp) and (not req and rsp);
eq policy : bool = today or yesterday;
eq since : bool =  alrt or (alrt and (false -> pre(since)));

guarantee "Start includes a waypoint" :
    rsp => wp;
guarantee "Locations required after the start waypoint" :
    (wp and not rsp) => avl;
guarantee "Waypoint is well-formed" :
    wp => WELL_FORMED_WAYPOINT(Waypoint);
guarantee "Alert if start is not bounded relative to a request" :
    policy or since;
\end{lstlisting}
\end{lrbox}

@@

The contract in \figref{fig:sw} uses \texttt{eq} statements to define variables local to the contract. For example, the \texttt{req} variable is equivalent to the \texttt{event(req)} expression. There is an implicit \emph{event} input (or output) associated with every named event port in a component. The semantics used here do not not buffer events so the implicit input (or output) is a boolean value. An \texttt{event} expression refers to that implicit input (or output) and is true if there is data on the named port. The system contract assumes well-formed input and guarantees a set of properties about the output.

@@

The guarantee is an invariant on the expression \texttt{policy or since} meaning that either the policy holds or the alert is sounding. The \texttt{policy} is satisfied if there is a request with a response or nothing in the current time step (\texttt{today}) or a request yesterday with its response today. The value of \texttt{yesterday} in the current time step relies on values from the previous time step. The \texttt{yesterday} value uses the \texttt{->} operator for initialization. The left operand is the initial value of \texttt{yesterday} at the start of the system, which in this example is \texttt{(req and not rsp)} because seeing a request with no response is inconclusive at the start.  The right operand is the value of \texttt{yesterday} after the initial step. Here the \texttt{pre} operator refers to the value of the expression \texttt{(req and not rsp)} in the prior time step. Intuitively, \texttt{yesterday} is true if the previous time step made a request without a matching response and the current time step has the matching response to that request with no new request.

The value of \texttt{since} relies on its own value in the previous time step. The intuitive reading of the expression is that the alert has been true since the time when it first sounded. The first \texttt{alrt} sets \texttt{since} to true, and once the value of \texttt{since} is true, that value persists as long as \texttt{alrt} holds. The \emph{Alert if start is not bounded relative to a request} guarantee defines one behavior of a correctly implemented system. Together with the other guarantees, the contract models the expected input and output of the system as a whole. Contracts also model the expected behavior of each component used to implement the system.

Model checking proves whether the contract models of the components, with their connections, implement the system model in \figref{fig:sw}. The results of the model checker are in \figref{fig:example}(b). The red exclamation points are properties that do not hold. Each comes with a counter-example. The result is not unexpected since the model of the route planner leaves the output behavior unconstrained (i.e., an untrusted component can do anything). The counter-example showing each vio with a counterexample that violates the policy and does not raise the alert.

@@

\newsavebox{\mntr}
\begin{lrbox}{\mntr}
\begin{lstlisting}[style=agree]
const is_latched : bool = 
  Get_Property(this, CASE_Properties::Monitor_Latched);
eq rsp : bool = event(Response);
eq req : bool = event(Request);
eq today : bool = (req = rsp);
eq yesterday : bool = (req and not rsp) ->
                      pre(req and not rsp) and (not req and rsp);
eq policy : bool = today or yesterday;
eq alert : bool = (not policy) 
               -> ((is_latched and pre(alert)) or not policy);
guarantee Monitor_Alert 
  "Alert port tracks alert variable" :
  event(Alert) = alert;
guarantee Monitor_Output
  "Output if not alerted" :
  if event(Alert) then (not event(Output)) else
  if event(Response) then (event(Output) and (Output = Response))
  else (not event(Output));
\end{lstlisting}
\end{lrbox}
